---
title: "ANNOVA and Regression"
author: "Eyob_Mulugeta"
date: "`r Sys.Date()`"
output:
  word_document: default
  word document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
tb=read.csv('https://raw.githubusercontent.com/Jobmrtall/MSD/refs/heads/main/tb_renamed.csv')
```

## One-Way ANOVA

We perform an ANOVA test to determine if the mean TB prevalence differs across regions.

```{r anova}
# Perform one-way ANOVA
aov_model <- aov(TB_Prev_100k ~ Region, data = tb)

# Display ANOVA summary
summary(aov_model)
```

## Assumption Checks for ANOVA

### 1. Normality of Residuals

We check if the residuals follow a normal distribution using the Shapiro-Wilk test and a Q-Q plot.

```{r normality}
# Check sample size and choose appropriate normality test
res <- residuals(aov_model)

if (length(res) >= 3 & length(res) <= 5000) {
  shapiro.test(res)  # Shapiro-Wilk test
} else if (length(res) > 5000) {
  library(nortest)
  ad.test(res)  # Anderson-Darling test for large datasets
} else {
  print("Sample size too small for normality test.")
}

```

The Shapiro-Wilk test helps determine if the residuals are normally distributed. A p-value greater than 0.05 suggests normality.

```{r qqplot}
# Q-Q plot for residuals
qqnorm(residuals(aov_model))
qqline(residuals(aov_model), col = "red")
```

A Q-Q plot is used to visually assess normality. If the points lie along the red reference line, normality is assumed.

### 2. Homogeneity of Variances

We use Levene’s test to check if the variance across groups is equal.

```{r homogeneity}
library(car)  # Load the 'car' package for Levene's test
leveneTest(TB_Prev_100k ~ Region, data = tb)
```

Levene’s test assesses the equality of variances. A p-value greater than 0.05 indicates homogeneity of variance.

### 3. Residuals vs. Fitted Values Plot

We plot the residuals against fitted values to check for homoscedasticity.

```{r residuals_plot}
plot(aov_model, which = 1)
```

This plot helps assess whether residuals are randomly scattered around zero, which is an assumption of ANOVA.

## Multiple Linear Regression

We perform a multiple linear regression to predict TB prevalence based on multiple independent variables.

```{r regression}
# Multiple linear regression model
multi_model <- lm(
  TB_Prev_100k ~ Year + Region + TB_Inc_100k + 
    TB_HIV_Inc_100k + TB_Mort_100k,
  data = tb
)
```

The linear model includes predictors such as Year, Region, TB Incidence, HIV percentage in incident TB cases, and TB Mortality to predict TB prevalence.

```{r regression_summary}
# Display summary of the regression model
summary(multi_model)
```

The summary provides estimates, standard errors, t-values, p-values, and R-squared values to evaluate model performance.
